---
title: "CF_CTRL"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

In this notebook. only include the analysis of the whole data after filtering, and the integration part. 

```{r,message = FALSE}
#REPO code for CF_seurat analysis
library(dplyr)
library(Seurat)
library(patchwork)
library(data.table)
library(R.utils)
library(ggplot2)
library(patchwork)
set.seed(412)
```

## Read the data

```{r,message = FALSE}
#load gene matrix
mat<-Seurat::ReadMtx("./CF_data/GSE145360_Sputum.data.processed.mtx",cells="./CF_data/GSE145360_Sputum.processed.barcodes.tsv",
                     features="./CF_data/GSE145360_Sputum.processed.genes.tsv", feature.column = 1)
#load metadata
meta <- data.frame(fread("https://cells.ucsc.edu/human-sputum/meta.tsv"), row.names=1)

```

## Data to Seurat object
```{r}
# Initialize the Seurat object with the raw (non-normalized data).
#New seurat object for paper analysis
cf <- CreateSeuratObject(counts = mat, project = "cf", min.cells = 3, min.features = 200,
                           meta.data=meta)

```

## QC 
```{r}
# 1. QC -------
Idents(cf) <- cf@meta.data$disease.ident
# get percent mitochondrial genes
cf[["percent.mt"]] <- PercentageFeatureSet(cf, pattern = "^MT-")
# Visualize QC metrics as a violin plot
VlnPlot(cf, features = c("nFeature_RNA", "nCount_RNA", "percent.mt")
        #, ncol = 3
        ,group.by = NULL
        )

plot1 <- FeatureScatter(cf, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(cf, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")+geom_smooth(method = 'lm')
plot1 + plot2

```


## data filtering
```{r}
# data filtering -----------------
cf.1 <- subset(cf, subset = nFeature_RNA>200 & nFeature_RNA < 7000 & percent.mt <5)

# split object
cf.1[["RNA"]] <- split(cf.1[["RNA"]], f = cf.1$disease.ident)
```

## normalize data
```{r}
# normalize data
cf.1 <- NormalizeData(cf.1)

# feature selection(find highly variable features)
cf.1 <- FindVariableFeatures(cf.1, selection.method = "vst", nfeatures = 3000)
# 
# ## Identify the 10 most highly variable genes
# top10 <- head(VariableFeatures(cf.1), 10)
# 
# ## plot variable features with and without labels
# plot1 <- VariableFeaturePlot(cf.1)
# plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
# plot1 + plot2
```
## Data Scale
```{r}
# scale to mean 0 and variance 1
all.genes <- rownames(cf.1)
cf.1<-ScaleData(cf.1, features=all.genes, verbose=F)
# regress out percentage number of mitochondrial genes
cf.1 <-ScaleData(cf.1, vars.to.regress = c("nFeature_RNA", "nCount_RNA","percent.mt"), do.scale=F, do.center=F, verbose=F)
```


## PCA dimensionaily reduction on subset of scaled variable features
```{r}
#PCA dimensionaily reduction on subset of scaled variable features
cf.1 <- RunPCA(cf.1, features = VariableFeatures(object = cf.1))

#visualize reduced dimensions
VizDimLoadings(cf.1, dims = 1:10, reduction = "pca")

#dimension heat map(first dimension- primary sources for heterogen.)
#cells and features ordered by PCA scores
#DimHeatmap(cf.1, dims = 1:20, cells = 500, balanced = TRUE)

#Determine "dimensionality" of dataset (paper used 12)
ElbowPlot(cf.1, ndims=20, reduction="pca")
```





## Perform streamlined (one-line) integrative analysis
We use rpca here after comparing our results with CCA and Harmony.
```{r}
# Perform streamlined (one-line) integrative analysis--------------------
CF_integration.1 <- IntegrateLayers(
  object = cf.1, method = RPCAIntegration,
  orig.reduction = "pca", new.reduction = 'integrated.rpca',
  verbose = FALSE)


CF_integration.1 <- FindNeighbors(CF_integration.1, reduction = "integrated.rpca", dims = 1:15)
CF_integration.1 <- FindClusters(CF_integration.1, resolution =c(0.1,0.2,0.3,0.5), cluster.name = c("rpca_clusters_0.1","rpca_clusters_0.2","rpca_clusters_0.3","rpca_clusters_0.5"))
CF_integration.1 <- RunUMAP(CF_integration.1, reduction = "integrated.rpca", dims = 1:15, reduction.name = "umap.rpca")

```

```{r}

DimPlot(CF_integration.1, reduction = "umap.rpca"
        , group.by = c("rpca_clusters_0.1","rpca_clusters_0.2"
                       ,"rpca_clusters_0.3","rpca_clusters_0.5")
        ,label = T)
# 
# DimPlot(CF_integration.1, reduction = "umap.unintegrated"
#         , group.by = c("rpca_clusters_0.1","rpca_clusters_0.2"
#                        ,"rpca_clusters_0.3","rpca_clusters_0.5")
#         ,label = T)

```

## findConserved markers

```{r}
# findConserved markers (resolution=0.1)--------------------
CF_integration.1 <- JoinLayers(CF_integration.1)
Idents(CF_integration.1) <- CF_integration.1@meta.data$rpca_clusters_0.1

# cluster0
marker_r1_0 <- FindConservedMarkers(CF_integration.1, ident.1 = 0, grouping.var = "disease.ident", verbose = FALSE)
head(marker_r1_0,20)
#FCGR3B,CXCR2,ALPL -- PMN by paper
VlnPlot(CF_integration.1, features=c("CXCR4", "IGF2R","ALPL"))
VlnPlot(CF_integration.1, features=c("FCGR3B", "CXCR2","ALPL"))

```


```{r}
# MRC1 or MARCO typically detected on alvMΦs 
VlnPlot(CF_integration.1, features=c("MRC1", "MARCO"))
```
After several searches, we name each cluster and plot the named clusters as below:

```{r}
CF_integration.1 <- RenameIdents(CF_integration.1, `0` = 'PMN')
CF_integration.1 <- RenameIdents(CF_integration.1, `1` = 'AM2')
CF_integration.1 <- RenameIdents(CF_integration.1, `2` = 'M/Momo')# 
CF_integration.1 <- RenameIdents(CF_integration.1, `3` = 'epithelial')
CF_integration.1 <- RenameIdents(CF_integration.1, `4` = 'T/B')
CF_integration.1 <- RenameIdents(CF_integration.1, `5` = 'AM1')

```
```{r}
# plot_clustering1 <- DimPlot(CF_integration.1,reduction = c("umap.unintegrated")
#         , label.size = 4,label = T)
plot_clustering2 <- DimPlot(CF_integration.1,reduction = c("umap.rpca")
          , label.size = 4,label = T)
# plot_clustering1|
plot_clustering2

DimPlot(CF_integration.1,reduction = c("umap.rpca")
          ,split.by = 'disease.ident',, label.size = 4,label = T)
```

## findMarkers between CF vs. Control

```{r}
# findMarkers between conditions ---------------------
CF_integration.1$celltype.cnd <- paste0(CF_integration.1$rpca_clusters_0.1,'_', CF_integration.1$disease.ident)
#View(CF_integration.1@meta.data)
Idents(CF_integration.1) <- CF_integration.1$celltype.cnd
DimPlot(CF_integration.1, reduction = 'umap.rpca', label = TRUE)
```



```{r}
# find markers
b.interferon.response <- FindMarkers(CF_integration.1, ident.1 = '1_CF', ident.2 = '1_CTRL')

head(b.interferon.response)
```


```{r}
# plotting conserved features vs DE features between conditions
#head(marker_r1_1)
FeaturePlot(CF_integration.1, features = c('G0S2', 'PRKCB', 'ANO5','AC026369.3'), split.by = 'disease.ident', min.cutoff = 'q10')
```

